Model Search in Relational Domains

### 2016-10-19 learning.tex
comments in section 6

\section{Abstract Conflict Analysis}
A conflict analysis in a SAT solver can be seen as abductive 
reasoning where a singleton assignment $c$ is replaced by a partial 
assignment that is sufficient to infer $c$.  Thus, a conflict 
analysis is a generalization of an underapproximation of a set of 
countermodels~\cite{sas12,dhk2013-popl}, where a countermodel is a 
set of structures that do not satisfy a formula.  Since relational 
abstract domains have more complex structure than a partial 
assignments domain, so the lifting of {\em first-UIP} algorithm~\cite{uip} 
and clause learning to any arbitraty lattice structure is a non-trivial task.

@@

\paragraph {\textbf{Lifting UIP to relational domains}}
An UIP is a special node in the abstract conflict graph such that 
any path from the last decision node to the conflict node must pass 
through it.  Unlike a SAT solver, a UIP in ACDCL may involve meet irreducibles 
that involves the same variables.  The first-UIP is a node closest to the 
conflict, and the last UIP is the decision node itself.  Computing UIPs ensures 
asserting cuts, that is, it yields clauses that generates new deduction after 
backtracking.  Every cut in the graph reason for conflict that can be used in learning.  
A first-UIP algorithm traverses the trail $\mathcal{T}$ starting from the 
conflict node and iteratively finds new nodes that contradict with the 
transformer stored in $reasons[\bot]$.

For our example, there exists multiple incomaparable reasons for conflict,
marked as {\em cut0, cut1}, in figure~\ref{conflict}.  Here, cut0 is the first UIP.  
Choosing cut0 yields a learnt clause 
$L_0: (x+y<15 \vee -x-y<-15 \vee -x+y<5 \vee x-y<-5 \vee y-10<0 \vee -y+10<0)$, 
which is obtained by negating the reason for conflict.  
%
\begin{figure}
%\caption{An example of Octagon}\label{octagon}
\scalebox{.65}{\import{figures/}{conflict_graph.pspdftex}}
\label{conflict}
\end{figure} 
%    
\paragraph{\textbf{Learning in Relational Lattice}}
We model learning as a meet irreducible deduction transformer, called 
{\em abstract unit transformer (AUnit)}.  For an 
abstract lattice $A$ with complementable meet irreducibles 
and a set of conflicting meet irreducibles (countermodels) 
$C \subseteq A$ of $\varphi$ such that \meet{C} does not 
satisfy $\varphi$, $AUnit_C: A \rightarrow A$ is defined as follows. 
\[ AUnit(a) =
 \begin{cases}
  \bot       & \quad \text{if } a \sqsubseteq \meet{C} \\
  \bar{t}    & \quad \text{if } t \in C \; \text{and} \; \forall t' \in C
  \setminus \{t\}. a  \sqsubseteq t' \\
  \bar{t}    & \quad \text{if } t \in C \; \text{and} \; \forall t' \in C \setminus \{t\}. a
  \not\sqsubseteq t', \; \text{then} \\ 
             & \quad \forall a' \subseteq decomp(a) \;  \text{and} \; \forall a'' \in a \setminus \{a'\}, \\ 
             & \quad \meet{a''} \sqsubseteq r' \; \text{and} \; a' \not\sqsubseteq t  \\
  \top & \quad \text{otherwise} \\
 \end{cases}
\]
For example, let $C = \langle x+z \geq 10,  k \geq 0 \rangle$ be the conflicting set 
of meet irreducibles of $\varphi$ and let $a= \langle x+y \geq 2, y+z \geq 5,
x+z \leq 15, k:[1,1] \rangle$.  Applying the third rule of $AUnit_C$, we get a 
decomposition of $a$ into $a'$ and $a''$ where 
$a'= \langle x+y \geq 2, y+z \geq 5, x+z \leq 15 \rangle \not\sqsubseteq (x+z \geq 10)$, and 
$a''= \langle k:[1,1] \rangle \sqsubseteq (k \geq 0)$.  Then, $AUnit_C(a) = (x+z < 10)$, 
which constrains the bound of the octagonal constraint.  
Let us consider another example, where $C = \langle x \geq 2, x \leq 5, y \leq 7 \rangle$
and let $a = \langle x:[3,4], y:[5,6] \rangle$, then $AUnit_C(a) = \bot$ using
the first rule, since $a \sqsubseteq C$.  Now, let $a' = \langle x:[3,4] \rangle$, 
then $AUnit_C(a') = (y > 7)$ using second rule, since $a' \sqsubseteq (x \geq 2)$
and $a' \sqsubseteq (x \leq 5)$.

@@

The abstract clause learning and backjumping procedures in the abstract 
conflict graph is stated in terms of the state of ACDCL solver as follows. 
\[AbsLearn: \quad  (\mathcal{E},S) \rightarrow (\mathcal{E},S \wedge
\mathcal{L}) \quad \text{if} \; \mathcal{L} \notin S \; \textrm{and}
\; (S \wedge \mathcal{L}) \; \text{is not UNSAT} \]
\[AbsBackjump: \quad (\mathcal{E}_1(m,s)\mathcal{E}_2,S) \rightarrow
(\mathcal{E}_1,S) \quad \text{if} \; (\mathcal{E}_1,S) \; \text{is
consistent} \]

>>>>

\section{Abstract Conflict Analysis}
A conflict analysis in a SAT solver can be seen as abductive 
reasoning where a singleton assignment $c$ is replaced by a partial 
assignment that is sufficient to infer $c$ \pscmt{???}.  Thus, a conflict 
analysis is a generalization of an under-approximation of a set of 
countermodels~\cite{sas12,dhk2013-popl}, where a countermodel is a 
set of structures that do not satisfy a formula \pscmt{this explanation does not help the reader, who will mix up counterexamples and countermodels}.  Since relational 
abstract domains have more complex structure than a partial 
assignments domain, so the lifting of {\em first-UIP} algorithm~\cite{uip} 
and clause learning to any arbitrary lattice structure is a non-trivial task.

@@

\paragraph {\textbf{Lifting UIP to relational domains}}
An UIP is a special node in the abstract conflict graph such that 
any path from the last decision node to the conflict node must pass 
through it.  Unlike a SAT solver, a UIP in ACDCL may have meet irreducibles 
that involve the same variables.  The first-UIP is a node closest \pscmt{unique?} to the 
conflict, and the last UIP is the decision node itself.  Computing UIPs ensures 
asserting cuts, that is, it yields clauses that generates new deduction after 
backtracking.  Every cut in the graph reason for conflict that can be used in learning.  
A first-UIP algorithm traverses the trail $\mathcal{T}$ starting from the 
conflict node and iteratively finds new nodes that contradict \pscmt{we have to define what that means in the preliminaries} with the 
transformer stored in $reasons[\bot]$.

For our example, there exists multiple incomaparable reasons for conflict,
marked as {\em cut0, cut1}, in figure~\ref{conflict}.  Here, cut0 is the first UIP.  
Choosing cut0 yields a learnt clause \pscmt{use $\leq$ instead of $<$} 
$L_0: (x+y<15 \vee -x-y<-15 \vee -x+y<5 \vee x-y<-5 \vee y-10<0 \vee -y+10<0)$, 
which is obtained by negating the reason for conflict.  
%
\begin{figure}[t]
%\caption{An example of Octagon}\label{octagon}
\scalebox{.65}{\import{figures/}{conflict_graph.pspdftex}}
\caption{\label{conflict}
\pscmt{push constants to the lhs of the inequalities; labels cut0, cut1 not fully visible}
}
\end{figure} 
%    
\paragraph{\textbf{Learning in Relational Lattice}}
We model learning as a meet irreducible deduction transformer \pscmt{???}, called 
{\em abstract unit transformer (AUnit)} \pscmt{why the name? seems confusing as we are talking about learning and not about deduction}.  For an 
abstract lattice $A$ with complementable meet irreducibles 
and a set of conflicting meet irreducibles (countermodels) \pscmt{The term countermodels for $C$ is misleading since $\bigsqcap C$ is a set of countermodels, but the elements of $C$ individually are no countermodels.}
$C \subseteq A$ of $\varphi$ such that $\bigsqcap C$ does not 
satisfy $\varphi$, $AUnit_C: A \rightarrow A$ is defined as follows. 
\pscmt{explain which cases are different in non-relational domains}
\[ AUnit(a) =
 \begin{cases}
  \bot       & \quad \text{if } a \sqsubseteq \meet{C} \\
  \bar{t}    & \quad \text{if } t \in C \; \text{and} \; \forall t' \in C
  \setminus \{t\}. a  \sqsubseteq t' \\
  \bar{t}    & \quad \text{if } t \in C \; \text{and} \; \forall t' \in C \setminus \{t\}. a
  \not\sqsubseteq t', \; \text{then} \\ 
             & \quad \forall a' \subseteq decomp(a) \;  \text{and} \; \forall a'' \in a \setminus \{a'\}, \\ 
             & \quad \meet{a''} \sqsubseteq r' \; \text{and} \; a' \not\sqsubseteq t  \\
  \top & \quad \text{otherwise} \\
 \end{cases}
\]
For example, let $C = \langle x+z \geq 10,  k \geq 0 \rangle$ \pscmt{why suddenly a different notation for $k$? The variable name $k$ might be confusing because it is usually used for counting or indices} be the conflicting set 
of meet irreducibles of $\varphi$ and let $a= \langle x+y \geq 2, y+z \geq 5,
x+z \leq 15, k:[1,1] \rangle$.  Applying the third rule of $AUnit_C$, we get a 
decomposition of $a$ into $a'$ and $a''$ where 
$a'= \langle x+y \geq 2, y+z \geq 5, x+z \leq 15 \rangle \not\sqsubseteq (x+z \geq 10)$, and 
$a''= \langle k:[1,1] \rangle \sqsubseteq (k \geq 0)$.  Then, $AUnit_C(a) = (x+z < 10)$, 
which constrains the bound of the octagonal constraint.  
Let us consider another example, where $C = \langle x \geq 2, x \leq 5, y \leq 7 \rangle$
and let $a = \langle x:[3,4], y:[5,6] \rangle$, then $AUnit_C(a) = \bot$ using
the first rule, since $a \sqsubseteq C$.  Now, let $a' = \langle x:[3,4] \rangle$, 
then $AUnit_C(a') = (y > 7)$ using second rule, since $a' \sqsubseteq (x \geq 2)$
and $a' \sqsubseteq (x \leq 5)$.

@@

The abstract clause learning and backjumping procedures in the abstract 
conflict graph is stated in terms of the state of ACDCL solver as follows. 
\[AbsLearn: \quad  (\mathcal{E},S) \rightarrow (\mathcal{E},S \wedge
\mathcal{L}) \quad \text{if} \; \mathcal{L} \notin S \; \textrm{and}
\; (S \wedge \mathcal{L}) \; \text{is not UNSAT} \]
\[AbsBackjump: \quad (\mathcal{E}_1(m,s)\mathcal{E}_2,S) \rightarrow
(\mathcal{E}_1,S) \quad \text{if} \; (\mathcal{E}_1,S) \; \text{is
consistent} \]   
\pscmt{What is $\mathcal{E}$? What is $m$? What is the notation $\mathcal{E}_1(m,s)\mathcal{E}_2$? What does consistent mean?}
\pscmt{Is there any difference between non-relational and relational?}
